---
title: "Homework 1: ggplot2"
author: "Kaisa Roots"
date: "2019-03-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

By using *mpg* dataset:

1. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

For continous variables, a gradient is shown in the legend section, however, only 2 variables are shown in the graph. Shape aesthetics can not be used for continous variables, however, shape can be mapped from a variable using *factor*. This shows that the year continous variable indeed consists of two variables - 1999 marked with a dot and 2008 marked with a triangle.  

- Color
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy, color = year))
```

- Size
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy, size = year))
```

- Shape

A continuous variable can not be mapped to shape

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy))
```

- Shape with factor
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy, shape = factor(year)))
```

2. What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy, color = year, size = year))
```

3. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

Stroke aesthetic can only be used for shapes that have a border (like 21, 24). 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy), shape = 24, color = "green" , fill = "#FFFF00", size = 5, stroke = 2)
```

4. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

In this case, car displacement values below 5 are shown in a different colour. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y = hwy, colour = displ < 5))
```
